Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-20 article/article.Rmd
Commit 2024-06-20 10:15

We studied the enhancement of the global horizontal irradiance (GHI), over Thessaloniki, 
for the period 1993 -- 2023, on one minute average records.
We identify the cloud enhancement events (CE), by creating an appropriate clear sky
irradiance reference with the use of a radiation transfer model, and aerosol optical 
depth (AOD) data, from the AERONET, and a collocated Brewer photospectrometer.
We found that there is a trend of CE cases of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
and a
mean total energy change, of the CE events by
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$
with the peak of the CE events are observed during May and June.
An analysis of the total duration of CE events, showed that durations longer than 
5 minutes are very rare, with exceptions lasting over an hour.
We have detected enhancements over the total solar irradiance, without exceeding the
maximum values reported for other sites, with more favorable CE conditions.

@@

An important aspect of the variability of solar radiation is its interaction with the
clouds. In general, clouds can attenuate a fraction  of solar irradiance, but under
certain conditions, can lead to  enhancement of the global horizontal irradiance
(GHI) reaching the ground. This effect can locally increase the observed GHI to
levels even higher than the expected clear-sky irradiance [@Cordero2023;
@Vamvakas2020; @CastillejoCuberos2020; @Vamvakas2020 and references therein].

Some of the proposed underling mechanisms of those enhancements, have been summarized
by @Gueymard2017; the most important being the scattering of radiation on the edges
of cumulus clouds.  It has been also suggested that enhancement of GHI can be
produced by thin cirrus clouds though refraction and scattering [@Thuillier2013].
Further investigation with radiative transfer modeling and observations pointed as
the prevailing mechanism, the strong forward Mie scattering through clouds of low
optical depth [@Pecenak2016; @Thuillier2013; @Yordanov2013; @Yordanov2015].  Overall,
enhancement event depend on different interactive factors, which include cloud
thickness, structure and type, and the relative position of the sun and the clouds
[@Gueymard2017; @Veerman2022].

On multiple sites, cloud enhancements (CE) have been reported to be able to briefly
exceed the value of the solar constant, resulting of clearness indices above unit. A
summary of extreme enchantment cases (ECE) has been compiled by @Martins2022. There
are also some practical implications of the cloud enchantments.  The intensity and
duration of enhancements can effect the efficiency and stability of photovoltaic
power production [@Lappalainen2020; @Jaervelae2020] while ECEs have the potential to
compromise the integrity of photovoltaic plants infrastructure [@DoNascimento2019].
It has also been demonstrated that these events can interfere in the comparison of
ground-based and satellite observations [@Damiani2018].  Global warming has likely
affected cloud coverage in the last few decades. @Liu2023 reported increases in cloud
cover over the tropical and subtropical oceans and decreases over most continents,
while @Dong2023 reported decreases over North America and Europe. To our knowledge,
there is no evidence on whether this trend has affected also the number and strength
of CE events.

>>>>

We studied the enhancement of the global horizontal irradiance induced by clouds over
Thessaloniki for the period 1993 -- 2023 using data recorded every one minute.  We
identified the cloud enhancement events (CE) by creating an appropriate clear sky
irradiance reference with the use of a radiative transfer model and aerosol optical
depth data from a collocated Cimel sun photometer and a Brewer spectrophotometer. We
found a trend in CE events of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
and a trend in the CE events irradiation of 
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June.  An analysis of the total
duration of CE events, showed that durations longer than 5 minutes are very rare,
with exceptions lasting over an hour.  Finally we have detected enhancements above
the total solar irradiance at the top of the atmosphere of up to
$400\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$200\,\text{W}/\text{m}^{2}$.

@@

An important aspect of the variability of solar radiation is its interaction with the
clouds. In general, clouds can attenuate a fraction  of solar irradiance, but under
certain conditions, can lead to  enhancement of the global horizontal irradiance
(GHI) reaching the ground. This cloud enhancement (CE) effect can locally increase
the observed GHI to levels even higher than the expected clear-sky irradiance
[@Cordero2023; @Vamvakas2020; @CastillejoCuberos2020; @Vamvakas2020 and references
therein].

Some of the proposed underling mechanisms of those enhancements, have been summarized
by @Gueymard2017; the most important being the scattering of radiation on the edges
of cumulus clouds.  It has been also suggested that enhancement of GHI can be
produced by thin cirrus clouds though refraction and scattering [@Thuillier2013].
Further investigation with radiative transfer modeling and observations pointed as
the prevailing mechanism, the strong forward Mie scattering through clouds of low
optical depth [@Pecenak2016; @Thuillier2013; @Yordanov2013; @Yordanov2015].  Overall,
enhancement events depend on different interactive factors, which include cloud
thickness, structure and type, and the relative position of the sun and the clouds
[@Gueymard2017; @Veerman2022].

On multiple sites, cloud enhancements have been reported that exceed momentarily the
solar constant, resulting in clearness indices above unit. A summary of extreme cloud
enchantment (ECE) cases has been compiled by @Martins2022. Cloud enhancements can
have  also some practical implications.  The intensity and duration of enhancements
can effect the efficiency and stability of photovoltaic power production
[@Lappalainen2020; @Jaervelae2020] while ECEs have the potential to compromise the
integrity of photovoltaic plants infrastructure [@DoNascimento2019].  It has also
been demonstrated that these events can interfere in the comparison of ground-based
and satellite observations [@Damiani2018].  Global warming has likely affected cloud
coverage in the last few decades. @Liu2023 reported increases in cloud cover over the
tropical and subtropical oceans and decreases over most continents, while @Dong2023
reported decreases over North America and Europe. To our knowledge, there is no
evidence on whether this trend has affected also the number and strength of CE
events.
